COFFEE: an Optimizing Compiler for Finite Element Local Assembly

### 2014-08-20 TACO-paper-v2/coffee.tex
Add TACO-paper-v2, which is the revised version

% Metadata Information
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2010}
\acmMonth{3}

@@

The time required to apply the numerical kernels is a major issue,
since the equation domain needs to be discretized into an extremely
large number of cells to obtain a satisfactory approximation
of the PDE, possibly of the order of trillions
\cite{Rossinelli2013}. For example, it has been well established that mesh
resolution is critical in the accuracy of numerical weather
forecasts. However, operational forecast centers have a strict time
limit in which to produce a forecast - 60 minutes in the case of the
UK Met Office. Producing efficient kernels has a direct scientific
payoff in higher resolution, and therefore more accurate,
forecasts. Computational cost is a dominant problem in
computational science simulations, especially for those based on
finite elements, which are the subject of this paper. We address, in
particular, the well-known problem of optimizing the local assembly
phase of the finite element method
\cite{Francis,quadrature1,petsc-integration-gpu,Kirby-FEM-opt}, which
can be responsible for a significant fraction of the overall computation
run-time, often in the range 30-60$\%$. With respect to these studies,
we propose a novel set of composable code transformations targeting, for
the first time, instruction-level parallelism - with emphasis on SIMD
vectorization - and register locality.

>>>>

% Metadata Information
%\acmVolume{9}
%\acmNumber{4}
%\acmArticle{39}
%\acmYear{2010}
%\acmMonth{3}

@@

The time required to apply the numerical kernels is a major issue,
since the equation domain needs to be discretized into an extremely
large number of cells to obtain a satisfactory approximation
of the PDE, possibly of the order of trillions, as in
\cite{Rossinelli2013}. For example, it has been well established that mesh
resolution is critical in the accuracy of numerical weather
forecasts. However, operational forecast centers have a strict time
limit in which to produce a forecast - 60 minutes in the case of the
UK Met Office. Producing efficient kernels has a direct scientific
payoff in higher resolution, and therefore more accurate,
forecasts. Computational cost is a dominant problem in
computational science simulations, especially for those based on
finite elements, which are the subject of this paper. We address, in
particular, the well-known problem of optimizing the local assembly
phase of the finite element method
\cite{Francis,quadrature1,petsc-integration-gpu,Kirby-FEM-opt}, which
can be responsible for a significant fraction of the overall computation
run-time, often in the range 30-60$\%$. With respect to these studies,
we propose a novel set of composable code transformations targeting, for
the first time, instruction-level parallelism - with emphasis on SIMD
vectorization - and register locality.
